\documentclass[a4paper,12pt]{article}
\usepackage{xcolor}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{geometry}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{titlesec}
\usepackage{tikz}
\usepackage{booktabs}
\usepackage{array}
\usetikzlibrary{shadows}
\usepackage{tcolorbox}
\usepackage{float}
\usepackage{lipsum}
\usepackage{mdframed}
\usepackage{pagecolor}
\usepackage{mathpazo}   % Palatino font (serif)
\usepackage{microtype}  % Better typography

\setlength{\parindent}{0pt}

% Page background color
\pagecolor{gray!10!white}

% Geometry settings
\geometry{margin=0.5in}
\pagestyle{fancy}
\fancyhf{}

% Fancy header and footer
\fancyhead[C]{\textbf{\color{blue!80}CS663 Assignment-5}}
% \fancyhead[R]{\color{blue!80}Saksham Rathi}
\fancyfoot[C]{\thepage}

% Custom Section Color and Format with Sans-serif font
\titleformat{\section}
{\sffamily\color{purple!90!black}\normalfont\Large\bfseries}
{\thesection}{1em}{}

% Custom subsection format
\titleformat{\subsection}
{\sffamily\color{cyan!80!black}\normalfont\large\bfseries}
{\thesubsection}{1em}{}

% Stylish Title with TikZ (Enhanced with gradient)
\newcommand{\cooltitle}[1]{%
\begin{tikzpicture}
\node[fill=blue!20,rounded corners=10pt,inner sep=12pt, drop shadow, top color=blue!50, bottom color=blue!30] (box)
{\Huge \bfseries \color{black} #1};
\end{tikzpicture}
}
\usepackage{float} % Add this package

\newenvironment{solution}[2][]{%
\begin{mdframed}[linecolor=blue!70!black, linewidth=2pt, roundcorner=10pt, backgroundcolor=yellow!10!white, skipabove=12pt, skipbelow=12pt]%
	\textbf{\large #2}
	\par\noindent\rule{\textwidth}{0.4pt}
}{
\end{mdframed}
}

% Document title
\title{\cooltitle{CS663 Assignment-5}}
\author{{\bf Saksham Rathi, Kavya Gupta, Shravan Srinivasa Raghavan} \\
\small Department of Computer Science, \\
Indian Institute of Technology Bombay \\}
\date{}

\begin{document}
\maketitle

\section*{Question 1}
\subsection*{(a) part}
\begin{solution}{Solution}
\textbf{\Large Procedure:-}

Let the translation operation for a $(x_0, y_0)$ displacement on an image $f_1$ be defined as $f_2(x, y) = f_1(x - x_0, y - y_0)$. Given images $f_1, f_2$ have size $N \times N$, their (Discrete) Fourier Transforms are related as $F_2(\mu, \nu) = F_1(\mu, \nu) e^{-j2\pi (\mu x_0 + \nu y_0)/N}$ (due to 2D-DFT Translation property).

As per Equation (3) of Section I of the paper, cross spectrum of $f_1$ and $f_2$ is:-
\begin{equation}
	C(\mu, \nu) = \frac{F_1(\mu, \nu) \, F_2^{*}(\mu, \nu)}{|F_1(\mu, \nu)||F_2(\mu, \nu)|} = e^{j2\pi (\mu x_0 + \nu y_0)/N}
	\label{eq:cross-spectrum}
\end{equation}

\hrule
\vspace{5pt}
\textbf{Proof for Equation \ref{eq:cross-spectrum} above:-}

Given $F_2(\mu, \nu) = F_1(\mu, \nu) e^{-j2\pi (\mu x_0 + \nu y_0)/N}$, we can write $F_2^{*}(\mu, \nu) = F_1^{*}(\mu, \nu) e^{j2\pi (\mu x_0 + \nu y_0)/N}$.
\begin{align*}
	\text{Now }|F_2(\mu, \nu)|^2 &= F_2^{*}(\mu, \nu) . F_2(\mu, \nu) \\
	&= F_1(\mu, \nu) e^{-j2\pi (\mu x_0 + \nu y_0)/N} . F_1^{*}(\mu, \nu) e^{j2\pi (\mu x_0 + \nu y_0)/N} \\
	&= F_1(\mu, \nu) . F_1^{*}(\mu, \nu) = |F_1(\mu, \nu)|^2
\end{align*}
Hence $|F_2(\mu, \nu)| = |F_1(\mu, \nu)|$. Substituting these values in Equation \ref{eq:cross-spectrum}, we get:-
\[C(\mu, \nu) = \frac{F_1(\mu, \nu) \, F_1^{*}(\mu, \nu) e^{j2\pi (\mu x_0 + \nu y_0)/N}}{|F_1(\mu, \nu)|^2} = e^{j2\pi (\mu x_0 + \nu y_0)/N}\]
\hrule
\vspace{5pt}
Let's take the Inverse (Discrete) Fourier Transform of $C(\mu, \nu)$ to get the autocorrelation function $c(x, y)$:-
\begin{align*}
	c(x, y) &= \mathcal{F}^{-1}\{C(\mu, \nu)\} = \mathcal{F}^{-1}\{e^{j2\pi (\mu x_0 + \nu y_0)/N}\} \\
	&= \frac{1}{N} \sum_{\mu=0}^{N-1} \sum_{\nu=0}^{N-1} e^{j2\pi (\mu x_0 + \nu y_0)/N} e^{j2\pi (\mu x + \nu y)/N} = \frac{1}{N} \sum_{\mu=0}^{N-1} \sum_{\nu=0}^{N-1} e^{j2\pi (\mu (x_0 + x) + \nu (y_0 + y))/N} \\
	&= \delta(x_0 + x, y_0 + y)
\end{align*}
See that $c(x,y)$ attains a non-zero value only at $(-x_0, -y_0)$, which is the displacement between the two images. Hence, the autocorrelation function of the cross spectrum is a delta function at the displacement $(-x_0, -y_0)$, hence helping us find value of $(x_0, y_0)$.

Hence the steps are:-
\begin{enumerate}
	\item Obtain 2D-DFT of $f_1$ and $f_2$. $f_2$'s can be found using the Fourier Shift Theorem.
	\item Compute the cross spectrum $C(\mu, \nu)$ using Equation \ref{eq:cross-spectrum}.
	\item Determine Inverse DFT of $C(\mu, \nu)$ and find its non-zero point. If it is at $(a, b)$, then the displacement vector is $(-a, -b)$.
\end{enumerate}

\textbf{\Large Time Complexity:-}

2D-DFT (IDFT) of an image of size $M \times N$ can be found using FFT efficiently in $O(MN \log MN)$ time. Hence for a $N \times N$ image, 2D-DFT (IDFT) takes $O(NN \log NN) = O(N^2 \log N)$ time.
\begin{itemize}
	\item Step 1 is 2D-DFT hence takes $O(N^2 \log N)$ time.
	\item Step 2 is a pointwise multiplication and division, hence takes $O(N^2)$ time.
	\item Step 3 is 2D-IDFT, hence takes $O(N^2 \log N)$ time.
\end{itemize}
Hence the total time complexity is $O(N^2 \log N)$. \\ \\
\textbf{\Large Comparison with Pixel-wise image comparison procedure:-}

Say there are $W_x$ values of x-axis translation shift to be checked and so is $W_y$ for y-axis. For each translation shift vector, we do this:-
\begin{enumerate}
	\item Create a new image by translating $f_1$ by $(x_0, y_0)$. This takes $O(N^2)$ time.
	\item Compare each pixel of the new image with $f_2$'s pixels (or take min squares difference). This takes $O(N^2)$ time.
\end{enumerate}
Hence the total time complexity is $O(W_x W_y N^2)$. This is much larger than $O(N^2 \log N)$ for $W_x, W_y \gg \log N$.
\end{solution}

\subsection*{(b) part}
\begin{solution}{Solution}
	As said in Section II of the paper, $f_2(x, y) = f_1(x \cos \theta_0 + y \sin \theta_0 - x_0, \, - x \sin \theta_0 + y \cos \theta_0 - y_0)$ ($f_2$ is rotated-then-translated version of $f_1$). We have to find rotation $\theta_0$ and shift $(x_0, y_0)$ values.
	\vspace{5pt}
	\hrule
	\vspace{5pt}
	\textbf{Lemma:} If 2D-FT of $f$ is $F(\mu, \nu)$, then 2D-FT of $f(x \cos \theta_0 + y \sin \theta_0, \, - x \sin \theta_0 + y \cos \theta_0)$ is $F(\mu \cos \theta_0 + \nu \sin \theta_0, \, - \mu \sin \theta_0 + \nu \cos \theta_0)$.

	\textbf{Proof:}
	\[
		F'(\mu, \nu) =\int_{\infty}^{\infty} \int_{\infty}^{\infty} f(x \cos \theta_0 + y \sin \theta_0, \, - x \sin \theta_0 + y \cos \theta_0) e^{-j2\pi (\mu x + \nu y)} \, dx \, dy
	\]
	Substitute $x' = x \cos \theta_0 + y \sin \theta_0$ and $y' = - x \sin \theta_0 + y \cos \theta_0$. We will get $x = x' \cos \theta_0 - y' \sin \theta_0$ and $y = x' \sin \theta_0 + y' \cos \theta_0$. Also,
	\[
		dx \, dy = |J| \, dx' \, dy' = \begin{vmatrix}
			\frac{\partial x}{\partial x'} & \frac{\partial x}{\partial y'} \\
			\frac{\partial y}{\partial x'} & \frac{\partial y}{\partial y'}
		\end{vmatrix} \, dx' \, dy' = \begin{vmatrix}
			\cos \theta_0 & -\sin \theta_0 \\
			\sin \theta_0 & \cos \theta_0
		\end{vmatrix} \, dx' \, dy' = dx' \, dy'
	\]
	Thus,
	\begin{align*}
		F'(\mu, \nu) &= \int_{\infty}^{\infty} \int_{\infty}^{\infty} f(x', y') e^{-j2\pi (\mu (x' \cos \theta_0 - y' \sin \theta_0) + \nu (x' \sin \theta_0 + y' \cos \theta_0))} \, dx' \, dy' \\
		&= F(\mu \cos \theta_0 + \nu \sin \theta_0, \, - \mu \sin \theta_0 + \nu \cos \theta_0)
	\end{align*}
	Hence proved.
	\vspace{5pt}
	\hrule
	\vspace{5pt}
	If $F_1(\mu, \nu)$ is the 2D-FT of $f_1$ and so is $F_2(\mu, \nu)$ of $f_2$, then:-
	\begin{align*}
		F_2(\mu, \nu) &= \mathbf{F} [f_1(x \cos \theta_0 + y \sin \theta_0 - x_0, \, - x \sin \theta_0 + y \cos \theta_0 - y_0)] \\
		&= e^{-j2\pi (\mu x_0 + \nu y_0)} \mathbf{F} [f_1(x \cos \theta_0 + y \sin \theta_0, \, - x \sin \theta_0 + y \cos \theta_0)] &\text{ (Fourier Shift Theorem)}\\
		&= e^{-j2\pi (\mu x_0 + \nu y_0)} F_1(\mu \cos \theta_0 + \nu \sin \theta_0, \, - \mu \sin \theta_0 + \nu \cos \theta_0) &\text{ (Lemma)}
	\end{align*}
	Observe that magnitudes of the FTs are equal (Say $M_1, \, M_2$ are magnitudes of $F_1, \, F_2$):-
	\begin{align*}
		M_2(\mu, \nu) = |F_2(\mu, \nu)| &= |e^{-j2\pi (\mu x_0 + \nu y_0)} F_1(\mu \cos \theta_0 + \nu \sin \theta_0, \, - \mu \sin \theta_0 + \nu \cos \theta_0)| \\
		&= |F_1(\mu \cos \theta_0 + \nu \sin \theta_0, \, - \mu \sin \theta_0 + \nu \cos \theta_0)| \\
		&= M_1(\mu \cos \theta_0 + \nu \sin \theta_0, \, - \mu \sin \theta_0 + \nu \cos \theta_0)
	\end{align*}
	Convert to polar coordinate system: $(\mu, \nu) \rightarrow (\rho, \theta)$, $\rho = \sqrt{\mu^2 + \nu^2}$, $\theta = \tan^{-1}(\nu/\mu)$. Then $M_2(\rho, \theta) = M_1(\rho, \theta - \theta_0)$. Hence this becomes like the problem in (a) part (but in 1D). Using the method in (a) part, we can find $\theta_0$.

	After the rotation angle $\theta_0$ is found, rotate $f_1$ by $\theta_0$ to get $f_3$. Now $f_3$ is related to $f_2$ by pure translation. Use the method in (a) part to find the translation vector $(x_0, y_0)$.

	Hence the steps are:-
	\begin{enumerate}
		\item Find 2D-FT of $f_1$ and $f_2$.
		\item Find the magnitudes of the FTs and convert to polar coordinates.
		\item Find the rotation angle $\theta_0$ using the method in (a) part (just in 1D).
		\item Rotate $f_1$ by $\theta_0$ to get $f_3$.
		\item Find the translation vector $(x_0, y_0)$ using the method in (a) part on $f_3$ and $f_2$.
	\end{enumerate}
\end{solution}
\end{document}